Cross Validation Approaches for Penalized Cox Regression

### 2018-05-05 cross-validation.tex
change figure 1,3, add definition of l_{i}

\section*{Abstract}
Cross validation is commonly used for selecting tuning parameters in penalized regression, but little research has been conducted to study the performance of cross validation in penalized Cox regression. Due to its partial likelihood construction, carrying out cross validation for Cox models is not straightforward, and there are several ways one could imagine applying it. Here, we propose two new cross-validation methods for Cox regression and compare them to approaches that have been proposed elsewhere.
\note{as well as to information criteria such as AIC and BIC -- I DON'T THINK THIS IS CURRENTLY IN THE PAPER ANYMORE?}
Our simulation studies show that in general, all cross-validation approaches tend to be conservative (i.e., select smaller models than the optimal choice of tuning parameters) for penalized Cox regression models. However, our proposed approach of cross-validating the linear predictors seems to offer an attractive balance of performance and numerical stability. We illustrate these advantages using simulated data as well as real data from two genome-wide studies of survival in cancer patients.
\input{main}

>>>>

\section*{Abstract}
Cross validation is commonly used for selecting tuning parameters in penalized regression, but little research has been conducted to study the performance of cross validation in penalized Cox regression. Due to its partial likelihood construction, carrying out cross validation for Cox models is not straightforward, and there are several ways one could imagine applying it. Here, we propose two new cross-validation methods for Cox regression and compare them to approaches that have been proposed elsewhere. Our simulation studies show that in general, all cross-validation approaches tend to be conservative (i.e., select smaller models than the optimal choice of tuning parameters) for penalized Cox regression models. However, our proposed approach of cross-validating the linear predictors seems to offer an attractive balance of performance and numerical stability. We illustrate these advantages using simulated data as well as real data from two genome-wide studies of survival in cancer patients.
\input{main}

### 2018-05-05 main.tex
change figure 1,3, add definition of l_{i}

In the Cox model \citep{Cox1975}, the hazard function for an observation i is given by \begin{equation}h_{i}(t) = h_{0}(t) \text{exp}( X_{i}^{T} \beta)\end{equation} where $h_{0}$ is the baseline hazard and $e^{X_{i}^{T} \beta}$ is the relative risk. Suppose there are n observations and m observed failure times, the estimation of the coefficients $\beta$s are obtained by maximizing the partial likelihood \begin{equation}L(\beta) = \prod_{j=1}^{m} \frac{\text{exp} ( X_{j}^{T} \beta)}{\sum_{ k \in R(t_{j})}\text{exp} ( X_{k}^{T} \beta)}\end{equation}, where j indexes the observed failure times and $R(t_{j})$ denotes the set of observations at risk at time $t_{j}$. As a semi-parametric model, Cox regression only gives estimations of the $\beta$ coefficients, without estimating the baseline hazard. Hence the interpretation of the Cox model is only valid in a relative sense.

@@

>>>>

Suppose there are n subjects and m observed failure events. Use i for indexing observations, j for indexing events sorted by their failure time. For the ith subject, its covarites will be denoted by $X_{i}$; $\delta_{i}$ denotes whether the event is observed or not for this subject. $t_{j}$ denotes the failure time for the j th event; $t_{i}$ denotes the time observed for ith subject.  In the Cox model \citep{Cox1975}, the hazard function for an observation i is given by 
\begin{equation}h_{i}(t) = h_{0}(t) \text{exp}( X_{i}^{T} \beta)\end{equation} 
where $h_{0}$ is the baseline hazard and $e^{X_{i}^{T} \beta}$ is the relative risk. The estimation of the coefficients $\beta$s are obtained by maximizing the partial likelihood. For each individual subject, its contribution to the partial likelihood is 
\begin{equation} l_{i}(\beta) = \left \{\frac{\text{exp} ( X_{i}^{T} \beta)}{\sum_{ k \in R(t_{i})}\text{exp} ( X_{k}^{T} \beta)}\right \}^{\delta_{i}} \end{equation} \note{introduce notation $l_{i}$} , where $R(t_{i})$ denotes the set of observations at risk at time $t_{i}$.

For all n subjects, the partial likelihood is 
\begin{equation} L(\beta) =\prod_{i = 1}^{n} l_{i}(\beta) = \prod_{j=1}^{m} \frac{\text{exp} ( X_{j}^{T} \beta)}{\sum_{ k \in R(t_{j})}\text{exp} ( X_{k}^{T} \beta)}\end{equation}. As a semi-parametric model, Cox regression only gives estimations of the $\beta$ coefficients, without estimating the baseline hazard. Hence the interpretation of the Cox model is only valid in a relative sense.

@@

\note{tranpose sign in figure 1 is corrected}
